### 2008-03-20 trunk/Papers/dl-gre-08/conclusion.tex
iterated 5-6

\section{Computing bisimulation classes} \label{sec:bisim}

>>>>

\section{Conclusion} \label{sec:conclusion}

In this paper, we have presented a new perspective on GRE as the
problem of computing concepts of description logic with given
extensions.  We have also shown how such concepts can be computed
efficiently by applying existing algorithms for computing bisimulation
classes.  Our algorithms are able to generate 95\% of the
non-relational and 67\% of the relational REs from
\newcite{viethen06:_algor_for_gener_refer_expres}.  They are also
extremely efficient (140 ms to generate REs for all individuals in a
random model of size 100); to our knowledge, these are by far the
fastest runtimes for relational GRE reported in the literature.  We
will make our implementations available online.

Because they compute referring expressions for all individuals in the
domain at once, our algorithms will perform especially strongly in
static settings, such as the generation of descriptions for museum
exhibits, in which the individuals and their properties don't change
much.  In such a setting, REs could be computed once when the
application starts, and then reused throughout.  Nevertheless, they
remain fast enough for real-time use in dynamic settings.  One
interesting question for future research is how bisimulation
algorithms that can incrementally update the bisimulation classes when
the model changes, such as \todo{cite}, can be adapted to GRE.

One finding in our experiments with the Viethen and Dale dataset is
that there is no single ordering that covers all human-produced
descriptions, which seems to be in contrast to Dale and Reiter's
\shortcite{Dale1995} assumption that there is only one ordering for
each given domain.  In fact, it is not even the case that each speaker
consistently uses just one ordering.  An interesting open research
question is thus what factors determine which ordering is used.
Unfortunately, both in the Viethen and Dale dataset and in the TUNA
corpus \cite{deemter06:_build_seman_trans_corpus_for}, only a minority
of referring expressions is relational, maybe because these domains
lend themselves very well to row/column style propositional REs.  We
are currently preparing an experiment to collect REs in a domain in
which relational REs will be more frequent.

### 2008-03-20 trunk/Papers/dl-gre-08/related.tex
iterated 5-6

\section{Related Work} \label{sec:related}

\cite{Dale1995}

\cite{deemter01:_gener_refer_expres}

\cite{dale91:_gener_refer_expres_invol_relat} -- avoid infinite
regress by requiring that each semantic atom can be used only once in
the description, i.e.\ no \emph{edge} may be used twice.

\cite{gardent02:_gener_minim_defin_descr}

\cite{Krahmer2003} -- avoid infinite regress automatically because
it doesn't count as progress to add an edge to the subgraph that was
already in it, i.e.\ cleaner version of Dale and Haddock heuristic.
Subgraphs can be seen as DL concepts made up of conjunction and
existential quantifiers (no negation).

\cite{kelleher06:_increm_gener_of_spatial_refer} -- generate GREs with
spatial relations. Polynomial algorithm (check this). Generates REs
for landmarks recursively, i.e.\ suffer from Dale and Haddock infinite
regress in principle. Avoid this by using each individual only once in
the entire description, i.e.\ no \emph{node} may be used
twice. Advantage over ours: take salience of landmarks into account.

>>>>

\section{A unified perspective on GRE} \label{sec:related}

\begin{figure}
  \centering
  \begin{tabular}{l|l}
    GRE algorithm & DL variant \\ \hline
    \newcite{Dale1995} & $\mathcal{CL}$ \\
    \newcite{deemter01:_gener_refer_expres} & $\mathcal{PL}$ \\
    \newcite{dale91:_gener_refer_expres_invol_relat} & \el \\
    \newcite{Krahmer2003} & \el \\
    \newcite{kelleher06:_increm_gener_of_spatial_refer} & \el \\
    \newcite{gardent02:_gener_minim_defin_descr} & $\mathcal{ELU}_{(\neg)}$
  \end{tabular}
  \caption{DL variants used by different GRE algorithms.}
  \label{fig:related}
\end{figure}

Viewing GRE as a problem of generating DL concepts offers a new,
unified perspective of the GRE problem: It is the problem of computing
a DL concept with a certain (typically singleton) extension.  Many
existing approaches can be subsumed under this view; we have
summarized this for some of them in Fig.~\ref{fig:related}, along with
the DL fragment they use.  We already discussed some of these
approaches in Section~\ref{sec:gre}.  Furthermore, the graphs
\newcite{Krahmer2003} use to represent referring expressions can be
seen as concepts of \el\ that are only satisfied at those points at
which they can be embedded; the non-relational but negative and
disjunctive descriptions generated by
\newcite{deemter01:_gener_refer_expres} are simply concepts of
$\mathcal{PL}$; and \newcite{gardent02:_gener_minim_defin_descr} then
generalizes this into generating concepts of $\mathcal{ELU}_{(\neg)}$,
i.e.\ \el\ plus disjunction and atomic negation.

The approach presented here fits very well into this landscape, and in
a sense completes the picture by showing how to generate concepts for
arbitrary bisimulation classes in \alc, which combines all connectives
used in any of these previous approaches.  Where our approach breaks
new ground is in the way these concepts are computed: It successively
refines a decomposition of the domain into subsets.  In this way, it
is reminiscent of the Incremental Algorithm, and in fact the IA can be
seen as the special case of the \el\ algorithm for the non-relational
case.

However, our approach is in contrast with
\newcite{dale91:_gener_refer_expres_invol_relat} and its successors,
such as \newcite{kelleher06:_increm_gener_of_spatial_refer}, which
attempt to generate a RE for a single individual for successive
individuals in the model and therefore run a risk of infinite regress.
Our algorithms are guaranteed to terminate (and with concepts of the
smallest relational depth that cover all the data), and infinite
regress is automatically excluded.  Perhaps closest in spirit is the
Krahmer et al.\ graph algorithm, which also computes \el\ concepts by
extending them successively; but this algorithm, too, is focused on
reference to a single individual, and it has NP-complete worst-case
complexity.
